---
title: "Medical Cost Analysis"
date: "2023-03-18"
output:
  distill::distill_article:
    toc: true
    toc_float: true
author: 
  - first_name: "Neal"
    last_name: "Carico"
    url: https://github.com/nealcarico/medical-cost-analysis.git
---
# Intro
> This exploratory data analysis was done with the skills I obtained from both the Google Data Analysis Certificate program as well as University of Michigan Data Analytics in the Public Sector Certificate Program.

### Preparing my environment and taking a peak at the data

```{r}
library(tidyverse) # metapackage of all tidyverse packages
data <- read.csv("data/insurance.csv")
```

```{r}
#Checking the data
summary(data)
head(data)
tail(data)
any(is.na(data))
```

## The Ask
> What insights can an insurance company get from this kind of data? 

### Questions and assumptions

  1. Age, BMI and smoking result in higher charges
  2. As long as the amount of data from each reason, it's safe to assume there won't be huge differences.
  3. Are there trends in the number of children and individual has?

## Processing
> This data was already cleaned prior to being published on Kaggle, and I just need some different forms of the metrics we already have.

```{r}
data$"has_children" <- ifelse(data$"children" > 0, TRUE, FALSE)
data$'smoker' <- ifelse(data$'smoker' == 'yes', TRUE, FALSE)
```

# Visualizations

### Age based charts

```{r}
data %>% 
  mutate(
    age_group = dplyr::case_when(
      age <= 18 ~ "0-18",
      age > 18 & age <= 29 ~ "19-35",
      age > 29 & age <= 49 ~ "30-49",
      age > 49 & age <= 59 ~ "50-59",
      age > 59 ~ "> 59"
    )) %>%
  ggplot(aes(x=age_group)) +
  geom_bar(fill="#69b3a2", color="#e9ecef", alpha=0.8)
```

```{r}
data %>%
  ggplot(aes(x=age, y=charges, color=region)) +
  geom_point(size=2)

```

### Smoking based charts

```{r}
data %>%
    ggplot(aes(x=smoker)) + 
    geom_bar(fill="#69b3a2", color="#e9ecef", alpha=0.8)
```

```{r}
data %>%
    ggplot(aes(x=age , y=charges, color=smoker)) + 
    geom_point(size=2)
```

### children related charts

```{r}
data %>%
    ggplot(aes(x=children)) + 
    geom_bar(fill="#69b3a2", color="#e9ecef", alpha=0.8)
```

```{r}
data %>%
    ggplot(aes(x=age, y=charges, color=has_children)) + 
    geom_point(size=2)
```

### Regional based charts

```{r}
data %>%
    ggplot(aes(x=region)) + 
    geom_bar(fill="#69b3a2", color="#e9ecef", alpha=0.8)
```

```{r}
data %>%
    ggplot(aes(x=age, y=charges, color=region)) + 
    geom_point(size=2)
```

### Sex based charts

```{r}
data %>%
  ggplot(aes(x=region)) +
  geom_bar(fill="#69b3a2", color="#e9ecef", alpha=0.8) + 
  facet_wrap(~sex)
```

```{r}
data %>%
  ggplot(aes(x=smoker)) +
  geom_bar(fill="#69b3a2", color="#e9ecef", alpha=0.8) + 
  facet_wrap(~sex)
```

```{r}
data %>% 
    ggplot(aes(x=age, y=charges, color=sex)) + 
    geom_point(size=2)
```

# EDA Conclusion

-   The biggest factor that increases the charges is if the individual is a smoker.
-   There is a small upward trend for age, not nearly as big of one as I expected.
-   BMI actually appears to have little effect on charges in this data set.
-   Region, sex, and having dependents on plan isn't a factor that increases charges.
-   There are 3 major trends that the data seems to follow that aren't quite identified yet (pictured in multiple plots including the last)
